---
title: "Group Histogram in R"
author: "Jarrett Cook"
date: "2025-05-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data description

The data used in this lab is from Sampledata2.csv. The dataset consists of 12 years of crime rate data for the United States. There are a total of 3 variables in the dataset, which are described below:

* Variable 1: State (all 50 States are included, from Alabama to Wyoming)
* Variable 2: Year (12 years of data are included, from 2007 to 2018)
* Variable 3: CrimeRate (crime rates are per 100,000 people)

# Data wrangling

Per the Module 6 assignment: Grouped Histogram in R, the following steps were taken with the provided code and output.  The printed result about the information on all the packages was hidden.

* Step 1: Load the required packages in R.
```{r, message=FALSE}
# Load tidyverse and colorspace libraries
library(tidyverse)
library(colorspace)
```
* Step 2: Set working directory
```{r}
setwd("C:/Users/jarco/OneDrive/Documents/DSCI605")
```
* Step 3: Load and inspect the dataset
```{r}
# Load and inspect the dataset
df <- read.csv("Sampledata2.csv")
view(df)
summary(df)
```
