---
title: "Group Histogram in R"
author: "Jarrett Cook"
date: "2025-05-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data description

The data used in this lab is from Sampledata2.csv. The dataset consists of 12 years of crime rate data for the United States. There are a total of 3 variables in the dataset, which are described below:

* Variable 1: State (all 50 States are included, from Alabama to Wyoming)
* Variable 2: Year (12 years of data are included, from 2007 to 2018)
* Variable 3: CrimeRate (crime rates are per 100,000 people)

# Data wrangling

Per the Module 6 assignment: Grouped Histogram in R, the following steps were taken with the provided code and output.  The printed result about the information on all the packages was hidden.

* Step 1: Load the required packages in R.
```{r, message=FALSE}
# Load tidyverse and colorspace libraries
library(tidyverse)
library(colorspace)
```
* Step 2: Set working directory.
```{r}
setwd("~/DSCI605")
```
* Step 3: Load and inspect the dataset. I used View() to inspect the dataset, but with 600 rows, I chose to show the head and tail along with the structure and summary.
```{r}
# Load and inspect the dataset
df <- read.csv("Sampledata2.csv")
str(df)
head(df)
tail(df)
summary(df)
```
* Step 4: Break the variable "CrimeRate" into three groups based on the crime rate range of <250, >500, or between 250 and 500 and put these groups under the new variable "RangeGroup."
```{r}
# Break the variable “CrimeRate” into three groups based on the range and create a new variable “RangeGroup”. You can use case_when() function.
df2 <- df %>% 
  mutate(RangeGroup = case_when(CrimeRate < 250 ~ "CrimeRate < 250",
                          250 <= CrimeRate & CrimeRate <= 500 ~ "250 <= CrimeRate <=500",
                          CrimeRate > 500 ~ "CrimeRate > 500"))

# Verify this worked by checking head() and tail(). Again, I used View(df2) but chose not to display all 600 rows.
head(df2)
tail(df2)
```
* Step 5: Create two color vectors.
```{r, echo=TRUE}
# Create the first color vector using colorspace.
n <- length(unique(df2$Year))
col1=colorspace::diverge_hcl(n)

# Create the second color vector using the colors green, blue, and red.
col2=c("green","blue","red")
```
* Step 6: Plot the first histogram.
```{r, echo=TRUE}
# Plot a histogram of the variable “CrimateRate” and fill with the variable “Year”. Use the first color vector col1 for this graph.
df2 %>% 
  ggplot(aes(x=CrimeRate, color=as.factor(Year), fill=as.factor(Year))) +
  geom_histogram(position = "dodge", alpha=0.5, bins=10) +
  scale_color_manual(values=col1) +
  scale_fill_manual(values=col1) +
  labs(title="Crime Rate Histogram Plot #1", x="Crime rate per 100,000 people", y="Count") +
  guides(col=guide_legend("Year"), fill=guide_legend("Year")) +
  theme_classic()
```

* Step 7: Plot the second histogram.
```{r, echo=TRUE}
# Plot a histogram of the variable “CrimeRate” and fill with the variable “RangeGroup”, which you created in Step3. Please use second color vector col2 for this graph.
df2 %>% 
  ggplot(aes(x=CrimeRate, color=as.factor(RangeGroup), fill=as.factor(RangeGroup))) +
  geom_histogram(position = "dodge", alpha=0.5, bins=10) +
  scale_color_manual(values=col2) +
  scale_fill_manual(values=col2) +
  labs(title="Crime Rate Histogram Plot #2", x="Crime rate per 100,000 people", y="Count") +
  guides(col=guide_legend("RangeGroup"), fill=guide_legend("RangeGroup")) +
  theme_classic()
```

